---
title: "Problem Set #1"
author: "Shuhan(Alice) Ai"
date: "09/30/2023"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 

Welcome to Introduction to Programming and Data Management using R! This problem set is intended to give you some practice becoming familiar with using R. In this problem set, we are asking you to: create an R project; knit to pdf; load and investigate an R data frame that is stored on the web; apply some basic functions to atomic vectors; and create a GitHub issue.

- Note: Change the values of the YAML header above to your name and the date. 


## Question 1: Creating an R project

**Create an R project**  
- Create a folder where you want to save files associated with problem set 1. Let's call that folder "problemset1", but you can name it whatever you want.
  - For instance, it could be EDUC260a >> problem_sets >> problemset1.
- In RStudio, click on "File" >> "New Project" >> "Existing Directory" >> "Browse".
- Browse to find and select your problem set 1 folder.
- Click on "Create Project".
  - An R project file has the extension ".Rproj".
  - The name of the file should be "problemset1.Rproj", or whatever you named the folder.

Save this problemset1.Rmd file anywhere in the folder named problemset1.

- Use this naming convention "lastname_firstname_ps#" for your .Rmd files (e.g. jaquette_ozan_ps1.Rmd).
  - If you want, you can change the name of this file to include your first and last name.
- Run the `getwd()` function and the `list.files()` function in the code chunk below.
- What is the output? Why?  

```{r, eval=FALSE}
getwd()
list.files()
```

**ANSWER:**

1. **Output**
   a. The output of the `getwd()` is `"/Users/aishuhan/Desktop/UCLA PhD/UCLA Course/EDUC 260A Data Management/problemset1"`
   b. The output of the `list.files()` function is `"ai_alice_ps1.Rmd"  "problemset1.Rproj"`. 
2. **Why**
   a. The output of the `getwd()` function is the path to the current working directory, which is where this R Markdown file is located
   b. The output of the `list.files()` function lists the files in the current working directory. In this case, it shows two files: "ai_alice_ps1.Rmd" and "problemset1.Rproj". 

## Question 2: Knit to pdf  

- At the top of this .Rmd file, type in your first and last name in the appropriate place in the YAML header (e.g. "Joe Bruin").
- In the date field of the YAML header, insert the date within quotations (any date format is fine).
- Now click the "Knit" button near the top of your RStudio window (icon with blue yarn ball) or drop down and select "Knit to PDF".
  - Alternatively you can use the shortcut: __Cmd/Ctrl + Shift + k__.

  - *Note*: One goal of this assignment is to make sure you are able to knit to a PDF without running into errors.

## Question 3: Load .Rdata directly with url and then investigate the data frame

1. This question asks you to load a dataframe by specifying the `url()` function within the `load()` function.
- Url link for data frame: https://github.com/anyone-can-cook/rclass1/raw/master/data/recruiting/recruit_ps1_small.Rdata

  - Hint: to load .Rdata use the `load()` and `url()` functions because you are using a link. follow this approach: `load(url("url_link"))`.
    - Note: the `url_link` is put within quotes.



Load the dataframe within this code chunk below.
```{r}
#?load
load(url("https://github.com/anyone-can-cook/rclass1/raw/master/data/recruiting/recruit_ps1_small.Rdata"))
```

2. Print the data frame `df_recruiting` by typing its name.
```{r}
df_recruiting
```

3. Use the `typeof()` function to investigate the type of data frame `df_recruiting`.
```{r}
typeof(df_recruiting)
```

4. Apply the `length()` function to the data frame `df_recruiting`. What does this output mean in your own words?
```{r}
length(df_recruiting)
```


**ANSWER:**

- It means that there are 6 elements in the data frame, which is also refers to the number of variables it contains. In this case, the `df_recruiting` has 6 variables, including 'state_code', 'school_type', 'ncessch', 'name', 'address', and 'city'.  

5. Use the `str()` function to investigate the structure of the data frame `df_recruiting`.
```{r}
str(df_recruiting)
```

6. Use the `names` function to list the names of the elements (variables) within `df_recruiting`.
```{r}
names(df_recruiting)
```

7. Wrap your answer above --- `names(data_frame_name)` --- within the `typeof()` function. Do the same for the `length()` function, and the `str()` function as well. Interpret what the output means in your own words.

```{r}
typeof(names(df_recruiting))
length(names(df_recruiting))
str(names(df_recruiting))
```

**ANSWER:** 

- `typeof(names(df_recruiting))`: This will return to the data type of the `name(df_recruiting)` function. The names of the elements within 'df_recruiting' are stored as character strings.
- `length(names(df_recruiting))`: This will return to the number of elements from `names(df_recruiting)`. In this case, there are 6 varibales. The length should be 6.
- `str(names(df_recruiting))`: The `str()` function aims to explore the data structure of `names(df_recruiting)`. In this case, the output refers to the structure of the character vector containing variable names.

## Question 4: Applying basic functions to atomic vectors

1. Create an atomic vector object named age with the following values: 3, 6, 41, 43.
```{r}
age <- c(3, 6, 41, 43)
```

2. Apply the `typeof()`, `length()`, and `str()` functions to the object `age`.
```{r}
typeof(age)
length(age)
str(age)
```

3. Apply the `sum()` function to `age`.
```{r}
sum(age)
```

4. Apply the `sum()` function to `age` but this time include the argument `na.rm = FALSE`.
```{r}
sum(age, na.rm=FALSE)
```

5. In general, what is a function "argument name" and what is an "argument value"? What does the argument `na.rm` do?

**ANSWER:** 

- In general, in a function call, an "argument name" is the identifier used to specify which parameter of the function you want to assign a value to, and an "argument value" is the actual value you assigned for that parameter.
- The `na.me` argument specifies whether to remove missing values (NA) when calculating the sum. When we set to `TRUE`, it tells R to remove missing valus before summing the vector, and when we set to `FALSE`, it includes missing values in the calculation. 

6. Create a new object `age2` with the following values: 3, 6, 41, 43, NA. Now calculate the sum of `age2` using the argument `na.rm = FALSE` and then calculate the sum using the argument `na.rm = TRUE`. Explain why the outputs of these two `sum()` functions differ.
```{r}
age2 <- c(3, 6, 41, 43, NA)
sum(age2, na.rm = FALSE)
sum(age2, na.rm =TRUE)
```

**ANSWER:** 

- `na.rm = TURE` tells R to ignore the missing value and only calculate the sum of the non-missing values, resulting in a sum of 93. 
- The outputs of these two `sum()` differ because the `na.rm` argument controls how missing values are operated during the calculation. When the argument value is set to `FALSE`, any presence of `NA` results in the entire sum being considered as `NA`. When the argument value is set to `TURE`, only the valid numeric values are included in the calculation.

7. Create a vector `tf` using the following code: `tf <- c(TRUE,FALSE,TRUE,FALSE,TRUE)`. Next apply the  `typeof()`, `length()`, and `str()` functions to the object `tf`. Based on this output, briefly describe the object `tf` in your own words (one sentence is fine).
```{r}
tf <- c(TRUE, FALSE, TRUE, FALSE, TRUE)
typeof(tf)
length(tf)
str(tf)
```


**ANSWER:** 

- The object `tf` is a logical vector consisting of five elements, where each element can either be `TURE` or `FALSE`.

8. Apply the `sum()` function to the object, using the option to remove `NA` values prior to calculation. What numeric value do mathematical calculations in `R` assign to `TRUE` values and what do they assign to `FALSE` values?
```{r}
sum(tf, na.rm = TRUE)
```

**ANSWER:** 

- `TRUE` is assigned the numeric value 1, and `FALSE` is assigned the numeric value 0. The result is 3.

9. This is the syntax of the `mean()` function that includes both argument names and the default values for arguments: `mean(x, trim = 0, na.rm = FALSE)`. 

When using a function, `R` requires you to type the values you assign to each argument, but typing in the argument names is usually optional. Even though it takes a bit more time, I usually like typing in both argument names and argument values, because it forces me to be more conscious about what value I am assigning to which argument, especially when a function is new to me. 

Use the `mean()` function to calculate the mean of object `tf` (removing `NA` values prior to calculation). In your function call, include both the argument name and the argument value for each argument (argument value for the `trim` argument can be `0`). Then run the same function, but without typing any argument names.

```{r}
#calculate the mean of 'tf' with explicit argument name and value
#trim=0 means no data points will be removed from the dataset
mean(x = tf, trim = 0, na.rm = TRUE)

#calculate the mean of 'tf' without specifying argument names
mean(tf, 0, TRUE)
```

## Question 5: Create a GitHub issue  


- Go to the [class repository](https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues) and create a new issue.

- Refer to [rclass1 student issues readme](https://github.com/anyone-can-cook/rclass1_student_issues_f23/blob/main/README.md) for instructions on how to post questions or things you've learned.

- You are also required to respond to at least one issue posted by another student.

- Paste the url to your issue here: https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues/93
- Paste the url to the issue you responded to here:
https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues/90

# Knit to pdf and submit problem set  

**Knit to pdf** by clicking the "Knit" button near the top of your RStudio window (icon with blue yarn ball) or drop down and select "Knit to PDF".

- Go to the [class website](https://anyone-can-cook.github.io/rclass1/) and under the "Readings & Assignments" >> "Week 1" tab, click on the "Problem set 1 submission link".  
- Submit both .Rmd and .pdf files.  
- Use this naming convention "lastname_firstname_ps#" for your .Rmd and pdf files (e.g. jaquette_ozan_ps1.Rmd & jaquette_ozan_ps1.pdf).
