---
title: "Problem Set #2"
author: "Shuhan (Alice) Ai"
date: "10/06/2023"
urlcolor: blue
output: 
  pdf_document:
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", highlight = TRUE)
  #comment = "#>" makes it so results from a code chunk start with "#>"; default is "##"
```

# Overview

In this problem set, you will practice investigating and subsetting objects using the Base R approach. We are asking you to practice working with different object types such as atomic vectors, lists, and data frames. Learning how to work with different object types will be very useful in the coming weeks and will help you understand the underlying structure of any data you are working with. While this problem set is fairly short, we want you to become familiar with these concepts -- which are fundamental to data management and working with R.

# Question 1: Practice making changes to the YAML header

We recommend reading [R Markdown: The Definitive Guide section 3.3](https://bookdown.org/yihui/rmarkdown/pdf-document.html), then make the following changes to the YAML header of this Rmd:

1.  Add your name to the YAML header.\
2.  Add a table of contents to YAML header.\
3.  The table of contents should have "depth" of 2.\
4.  Add section numbering to headers.

# Question 2: Investigating objects, Base R

Run the following to load data frame objects:

```{r}
rm(list = ls()) # remove all objects

load(url("https://github.com/ozanj/rclass/raw/master/data/recruiting/recruit_school_allvars.RData"))
#glimpse(df_school_all)

dim(df_school_all)
```

1.  Answer the following questions about the object `df_school_all` by running the appropriate R command in the code chunks below and write your response below the question following the **ANSWER:** prompt. The first question will be answered for you to show how it works.

    -   What "type" of object is `df_school_all`?
        -   **ANSWER [GIVEN]: The object `df_school_all` has type equals `r typeof(df_school_all)`.**\
    -   What is the "length" of the object `df_school_all`? What does this specific value of length refer to?
        -   **ANSWER: The length of the object `df_school_all` equals `r length(df_school_all)`. It refers to the number of variables in the object `df_school_all`.**\
    -   How many "rows" are in the object `df_school_all`? What does each row represent?
        -   **ANSWER: There are `r nrow(df_school_all)` rows in the object `df_school_all`, 21301 refers to the number of observations or records in the data frame.**

```{r}
typeof(df_school_all)
length(df_school_all)
nrow(df_school_all)
```

2.  In the below code chunk, use the `str()` function to describe the contents of `df_school_all` and then answer the following questions:

    -   What does each element of the object `df_school_all` represent? (Hint: Lecture Investigating data patterns in Base R, Slide 17)
        -   **ANSWER: Each element refers to a variable associated with schools.**\
    -   Are the individual elements within `df_school_all` lists or vectors?
        -   **ANSWER: The elements in `df_school_all` are vectors. In this context, they are either character vectors (e.g., `state_code`, `school_type) or numeric vectors (e.g.,`ptc_white\`).** \
    -   Are the individual elements within `df_school_all` named or un-named? If named, what do these element names refer to?
        -   **ANSWER: The individual elements within `df_school_all` are named. The names correspond to the variable names in the dataset. For example, `state_code`, `school_type`, `name`, etc., are the names of the variables.**

```{r, results='hide'}
str(df_school_all)
```

3.  These questions refer to the variable `school_type` within the object `df_school_all`. For the first two questions, run the appropriate R command in the code chunk below and write your response below the question.

    -   What is the data "type" of `school_type`?
        -   **ANSWER: The data type of `school_type` is `r typeof(df_school_all$school_type)`.**\
    -   What is the "length" of `school_type`? What does this specific value of length refer to?
        -   **ANSWER: The length of `school_type` is `r length(df_school_all$school_type)`.**

```{r}
typeof(df_school_all$school_type)
length(df_school_all$school_type)
```

4.  In these questions, you will apply the `table()` function to the variable `school_type` within the object `df_school_all`.

    -   In your own words, what does the `table()` function do?
        -   **ANSWER: The `table()` function in R is used to generate a frequency table, which displays the counts of unique values in `school_type`.**\
    -   What does the `useNA` argument of the `table()` function control?
        -   **ANSWER: The `useNA` argument of the `table()` function controls whether or not missing values are included in the frequency table. **\
    -   What is the default value of the `useNA` argument and what does this default value mean?
        -   **ANSWER: The default value the `useNA` argument in the `table()` function is `"ifany"`. This means that if there are any missing values in the data, they will be included in the frequency table.**\
    -   What happens when you assign the value `"ifany"` to the `useNA` argument?
        -   **ANSWER: When I assign the value `"ifany"` to the `useNA` argument, missing values will be included in the frequency table only if they are present in the data. If there are no missing values, the frequency table will not include an entry for NA.**\
    -   What happens when you assign the value `"always"` to the `useNA` argument?
        -   **ANSWER: When I assign the value `"always"` to the `useNA`, it means that missing values will always be included in the frequency table as a seperate column, regardless of whether they are present in the data or not.**\
    -   In the below R code chunk, use the `table()` function to count the number of observations for each value `school_type` three different ways:
        -   First, without specifying any value for `useNA`\
        -   Second, by assigning the value `"ifany"` to the `useNA` argument\
        -   Third, by assigning the value `"always"` to the `useNA` argument

```{r}
table(df_school_all$school_type)
table(df_school_all$school_type, useNA = "ifany")
table(df_school_all$school_type, useNA = "always")
```

# Question 3: Subsetting, Base R

In the code chunk below, you will find 3 objects: a vector, a list, and the data frame `df_school_all`. Run the code chunk. You will use these objects in the following questions.

```{r}
# Create a named numeric atomic vector
vec <- c(a = 2.4, b = 1.1, c = 3.4, d = 4, e = 6, f = 32, g = 21, h = 17, i = 10)
str(vec)

# Create a list
list <- list(c(1:3), list("red", "orange"), list("LA", "NY", "DC")) 
str(list)

# View the `df_school_all` data frame you loaded earlier
head(df_school_all, n = 5)
#str(df_school_all)
```

1.  In this question we will use the `[]` to subset the atomic vector `vec`:

    -   Return the 4th and 7th element of the vector `vec` (Hint: We are subsetting elements by position)\
    -   Return everything but the last element of the vector `vec`\
    -   Return elements named "a", "d", and "g"\
    -   Return elements that are less than `12`

```{r}
vec[c(4, 7)]
vec[-length(vec)]
vec[c("a", "d", "g")]
vec[vec < 12]
```

2.  In this question we will use the `[]` to subset columns in `df_school_all`:

    -   Return the first 100 elements of the `state_code` column\
    -   Return all addresses in the `address` column that are more than 30 characters long (Hint: Use `nchar()` to get number of characters, you can type ?nchar in the console or code chunk to get more information on the function)\
    -   Return all cities in the `city` column that are either 21 or 25 characters in length

```{r}
df_school_all$state_code[1:100]
df_school_all$address[nchar(df_school_all$address) > 30]
df_school_all$city[nchar(df_school_all$city) %in% c(21, 25)]
```

3.  In this question we will use the `[]` to subset a list/data frame:

    -   Return the 1st element of the list `list` using `[]`
        -   What is the data type? **ANSWER: The data type is "list". Using `[]` with a list always returns a list. **\
    -   Return the 2nd, 4th, and 6th elements of the data frame `df_school_all`\
    -   Return the first 3 rows (observations) and the variable names `state_code` and `name` of the data frame `df_school_all`

```{r, results='hide'}
list[1]
typeof(list[1])
df_school_all[c(2, 4, 6)]
df_school_all[1:3, c("state_code", "name")]
```

4.  In this question we will use `[[]]` and `$` to subset a list/data frame:

    -   Return the 1st element of the list `list` using `[[]]`
        -   What is the data type? **ANSWER: The data type is `r typeof(list[[1]])`. This is because the first element of the list `list` is a numeric vector.**\
    -   Return the variable `total_students` using `$`. The output here will be too large when you try to knit to PDF at the end of the problem set so just makem this line of code a comment by putting a hashtag \# in front of it.
        -   What is the data type? **ANSWER: The data type is `r typeof(df_school_all$total_students)`. This is because `total_students` represented as numeric vectors.**

```{r}
list[[1]]
typeof(list[[1]])
#df_school_all$total_students
typeof(df_school_all$total_students)
```

# Question 4: Create a GitHub issue

-   Go to the [class repository](https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues) and create a new issue.

-   Refer to [rclass1 student issues readme](https://github.com/anyone-can-cook/rclass1_student_issues_f23/blob/main/README.md) for instructions on how to post questions or reflections.

-   You are also required to respond to at least one issue posted by another student.

-   Paste the url to your issue here:https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues/194

-   Paste the url to the issue you responded to here:https://github.com/anyone-can-cook/rclass1_student_issues_f23/issues/191

# Knit to pdf and submit problem set

**Knit to pdf** by clicking the "Knit" button near the top of your RStudio window (icon with blue yarn ball) or drop down and select "Knit to PDF"

-   Go to the [class website](https://anyone-can-cook.github.io/rclass1/) and under the "Readings & Assignments" \>\> "Week 2" tab, click on the "Problem set 2 submission link"\
-   Submit both .Rmd and pdf files\
-   Use this naming convention "lastname_firstname_ps#" for your .Rmd and pdf files (e.g. jaquette_ozan_ps2.Rmd & jaquette_ozan_ps2.pdf)
